\chapter*{Acknowledgments}
\labelchapter{acknoledgments}
%-----------------------
\addcontentsline{toc}{chapter}{Acknowledgments}
\beginchapter
%-----------------------
\par \noindent
The work accomplished throughout this Bachelor Thesis and Project would not have been done the way it was, and the output achieved would not be the same without the help and support of many people and organizations. I would like to particularly thank everyone of the following for his/her help.

\vspace{0.4cm}

\par \noindent
My Father {\bf Mofid Boktor}, my mother {\bf Elham Fahim}, and my brother {\bf Peter Boktor} for their continuous support through the whole project period, their effort done to make my travel as organized and as easy as possible.

\vspace{0.4cm}

\par \noindent{\bf  Prof. Dr. Thomas Ertl} for accepting to supervise our project idea and giving us the opportunity of studying in the Universit\"at Stuttgart.

\vspace{0.4cm}

\par \noindent{\bf  Dipl. Inf. Christiane Taras} and {\bf Dipl. Inf. Steffen Koch}, being our co-supervisors, for their support, their advices and encouragement during the project and our stay in Stuttgart.

\vspace{0.4cm}

\par \noindent{\bf  Dr. Rimon Elias} for supervising our project in the German University in Cairo.

\vspace{0.4cm}

\par \noindent{\bf Prof. Dr. Paul J. K\"uhn} for helping us to come to Stuttgart to accomplish our Bachelor Thesis.

\vspace{0.4cm}

\par \noindent{\bf  Frederic-Gerald Morcos} and {\bf Marleine Daoud}, my team members, for their help and encouragement throughout the project.

\vspace{0.4cm}

\par \noindent{\bf  Abdel-Karim Mardini} and {\bf Karim Ratib}, for their sponsorship and the project idea: OpenGrafik.

\vspace{0.4cm}

\par \noindent{\bf  Youssef Abou-Kewik} for his sponsorship of OpenGrafik and for offering his and his company's {\bf connectmeTV} help.

\vspace{0.4cm}

\par \noindent{\bf Mohamed Essam} for his help in the MIE, IEEE competition.

\vspace{0.4cm}

\par \noindent

\vspace{0.4cm}